The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-06-23 Paper/jaccardbaetica.tex
updated maps

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.

@@

The strategic location of the workshops close to the river allowed the transport of olive oil through river connectivity to the routes of the Mediterranean and the Atlantic \citep{garcia_vargas_enrique_formal_2010}.

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/workshop1}
\caption{Distribution of the workshops (dots) in Baetica province}
\label{workshop}
\end{figure}

@@

The frequency of stamps per workshop can be seen in Fig. \ref{stamps}. Most 
workshops contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps. The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/britmap2.pdf}
\caption{Distribution of centres in \textit{Britannia}}
\label{britannia}
\end{figure}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/germania5_2.pdf}
\caption{Distribution of the sites in \textit{Germania}. Most sites were located in German \textit{limes}}
\label{germania}
\end{figure}

>>>>

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}.

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/general_map}
\caption{Overview of the sites analyzed in this paper. The color defines the three different regionsu under study (Baetica as producer; Germania and Britannia as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sitese}

\label{general}
\end{figure}

In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.

@@

The strategic location of the workshops at the riverbank allowed the transport of olive oil through riverine transport towards the maritime routes connecting both the Mediterranean and the Atlantic \citep{garcia_vargas_enrique_formal_2010}.

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/baetica}
\caption{Distribution of stamps found in Dressel-20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found on each site}

\label{workshop}
\end{figure}

@@

The frequency of stamps per workshop can be seen in Fig.~\ref{stamps}. Most 
workshops contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps. The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/britannia}
\caption{Sites in Britannia where Dressel-20 amphoric stamps have been found. Each dot is a site while its size shows the sample size of stamps found in the site. A majority of stamps have been found in garrisons related to Hadrian's and Antonine walls}
\label{britannia}
\end{figure}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/germania}
\caption{Sites in Germania where Dressel-20 amphoric stamps have been found. Each dot is a site while its size shows the sample size of stamps found in the site. Most sites are located along the German \textit{limes}}
\label{germania}
\end{figure}
